### 2021-03-26 chapters/sections/6-online_access_law.tex
finished

This chapter gives an overview about the Online Access Law regarding its regulations, their implementation, the basic OZG use case and the underlying system architecture.

@@

In 2017, the German government passed the Online Access Law (OZG) which requires federal republic and member states to execute the following regulations until 2022 \cite{ozg:regulations}:

\paragraph{Digital availability of administrative services} An administrative service is the electronic processing of administrative procedures that are available from outside the governmental institution.  As it is not clear which administrative services exactly are included by the definition of the OZG, the Federal Ministry of the Interior Building and Community (Bundesministerium des Innern, für Bau und Heimat - BMI) created a catalogue \cite{ozg:informationsplattform}. The OZG requires these services to be digitally available. As a guideline to what is considered sufficient availability, the BMI defined a maturity model \cite{ozg:maturity_model}.

\paragraph{Digital access to administrative services through administration portals of a portal network} Federal republic, each federal state and each commune must provide an administration portal. Portals of communes must be linked to the portal of the corresponding member state. Portals of federal republic and member states must be connected through a portal network. \cite{ozg:onlinegateway_general} Each portal must provide a "seek and find" feature, which enables users to find all administrative services provided by any administration portal \cite{ozg:onlinegateway_architecture}.

\paragraph{Interoperable user profiles for accessing administrative services} Federal republic and member states must provide user profiles that can be used to identify the corresponding person while requesting access to administrative services, to save personal information according to the once-only principle, to receive and send messages via a digital mailbox and to pay for services. The user profiles must be interoperable for every administration portal of the portal network \cite{ozg:user_profiles_general}.

@@

The implementation of OZG regulations can be separated into two projects: The digitalization and the networking project.

The goal of digitalization in this case is, to make administrative services available towards a user in a digitized way: A digitized administrative service can be accessed by a user through a website. The website is hosted either by the federal republic or a federal state and is called "Administration Portal". Access is usually provided through an application form. The administrative service can be managed through a user profile which is provided by the federal republic or the member state. Management of an administrative service usually includes starting the service by sending in a form, communicating with the responsible institutions through the inbox of the profile and receiving a result. The user profile can also enable users to upload documents to a datawallet and to save personal information for automatically filling in forms.

Digitalization of an administrative service is the modification of the underlying processes to incorporate the usage of the described features of the user profile and administration portal. In total, the BMI lists 575 relevant services, some of them provided by the federal republic, some by the member states and some by the communes \cite{BMI:Onlinezugangsgesetz}.

@@

In order to save investments a method called "one for all" is used when hosting administrative services. One federal state or the federal republic provides access to a service on their administration portal and distributes the requests to the responsible institutions "under the hood". As administration portals are connected through an "online gateway", each portal contains a search feature, which enables users to find all administrative services through any portal. Interoperable user profiles enable usage of each profile for management of administrative services on all portals.

@@

A user wants to apply for two administrative services: funding under the Federal Training Assistance Act (BAföG) and a drivers license. Both administrative services are provided through the "one-for-all" method. The federal state Sachsen-Anhalt is responsible for processing BAföG applications on its administration portal and the federal state Hessen is responsible for processing drivers license applications on its administration portal. The user has never access OZG services before.

\paragraph{Create User Profile} The user wants to apply for BAföG first and therefore visits an administration portal which informs them, that in order to apply for administrative services, a user profile is required. As user profiles are designed to be interoperable, it does not matter through which administration portal the user profile is created.

The user is requested to enter personal information and to specify an authentication method. For authentication for example a username and password combination can be used.

@@

\paragraph{Login to User Profile} Before being able to apply for administrative services, the user has to login into the user profile on the administration portal. Depending on the authentication method used for logging in, a trust level of the current session is determined. Only with a high trust level users can apply for a broad selection of administrative services. The user has selected an authentication method which results in a high trust level.

\paragraph{Application} On the web page of the administrative portal the user clicks on a "bring me to the application" button and is forwarded to a web page with an interactive form. Before the user is forwarded, he has the option to transfer personal information for automatic filling of the form. The user decides to accept and is presented with a filled in form. After pressing the "submit" button the application is transferred to the responsible institution.

\paragraph{Management of Applications} The user is forwarded back to the administration portal and selects to view all active applications created through the portal. Here the user can see the application they just submitted and is able manage it by for example requesting cancellation.

\paragraph{Communication} Through the inbox of the user profile the user is notified about a message regarding the application. The institution responsible for processing the application informs the user about receiving the application. Eventually the institution finishes processing the application and sends a message to the inbox of the user to inform them about the result. This marks the end of the basic OZG use case, however, the user wants to apply for a second administrative service.

\paragraph{Selection of Administrative Service} Using the search feature of the administration portal, the user now enters the search term "drivers license", selects the correct search result and is forwarded to the corresponding web page on the administration portal of Hessen.

@@

\paragraph{Application} The application process is designed to be similar for each administration portal, however, as Sachsen-Anhalt and Hessen operate their application processes independently, minor differences exist in for example the design and text of the web pages.

\paragraph{Management of Applications} The user is, again, forwarded back to the administration portal and selects to view all active applications created through the portal. The user is only able to see and manage the application for a drivers license. As the application for BAföG was created through a the administration portal of Sachsen-Anhalt, it is not accessible through the administration portal of Hessen.

@@

\subsection{System Architecture}
This section describes the OZG system architecture of a federal state capable of executing the basic OZG use case. References about OZG system architectures of Baden-Württemberg \cite{ozg:bw} and Nordrhein-Westfalen \cite{ozg:nrw} are used. The presentation document regarding the administration portal of Baden-Württemberg \cite{ozg:bw_administration_portal} is used as basis for describing the administration portal.

As shown in figure \ref{ozg:system_overview}, the OZG system architecture is separated into three components and two domains. Domains in this case separate systems in regard to the physical location where they operate. Components in this case describe functionally correlated systems of a system architecture.

The "User Component" directly interacts with users to enable access to administration portals. This can be desktop or mobile applications or as in this specific case a web browser. The "Administration Portal Component" hosts the administration portal. The "User Profile Component" provides interoperable user profiles. The "Integration Component" transfers filled in applications to responsible institutions. And finally, the "Institution Component" receives and processes applications.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{Diagrams/OZG System Overview.pdf}
    \caption{OZG System Overview}
    \label{ozg:system_overview}
\end{figure}

The administration portal directly interacts with the "User Component" through a web page and provides access to the basic OZG use case by interacting with other components. It consists of a web server that is responsible for hosting web pages and a service facade connected to multiple service components. Through the service facade, the web server accesses each service component and makes its service available to the "User Component". 
The "User Profile Management" component is capable of interacting with the "User Profile Component" to create new user profiles, retrieve personal information from existing user profiles and process login requests. 
The "User Profile Component" provides conventional user profile based identity management. 
Through the Federated Identity Management of Interoperable User Accounts in Germany (Föderiertes Identitätsmanagement interoperabler Nutzerkonten in Deutschland - FINK) the "User Profile" component of each federal state is made interoperable. The "Search" component can be used to retrieve a list of URLs of administrative services corresponding to a search term. It can retrieve URLs of administrative services from all administration portals through the "Online-Gateway".

The "Online-Gateway" is a system connecting all administration portals and enabling them to exchange information. The "Formula Engine" component is accessed by the web server to display interactive forms to the user when applying for selected administrative services. The interactive form can be either integrated into the web page of the administrative service or hosted on a separate web page. The user can opt in to automatically fill in forms with personal information. In that case, after authentication of the user, the "Administration Portal" component retrieves personal information from the "User Profile" component and sends it to the "Formula Engine" before displaying a filled in form to the user.

The "Application Management" component receives submitted applications by the "Formula Engine" component for further processing. Depending on the administration portal and type of application, additional processing steps, that are not part of the use case can take place. This includes for example payment processes. The component tracks the current status of the application which the user can view through the web page of the administration portal. It is also capable of terminating the application process if the user requests it.

>>>>

This chapter gives an overview about the Online Access Law regarding its regulations, their implementation, the basic OZG use case, and the underlying system architecture.

@@

In 2017, the German government passed the Online Access Law (OZG), which requires the federal republic and federal states to execute the following regulations until 2022 \cite{ozg:regulations}:

\paragraph{Digital availability of administrative services} An administrative service is the electronic processing of administrative procedures that are available from outside the governmental institution.  As it is not clear which administrative services exactly are included in the definition of the OZG, the Federal Ministry of the Interior Building and Community (Bundesministerium des Innern, für Bau und Heimat - BMI) created a catalogue \cite{ozg:informationsplattform}. The OZG requires these services to be digitally available. As a guideline on what is considered sufficient availability, the BMI defined a maturity model \cite{ozg:maturity_model}.

\paragraph{Digital access to administrative services through administration portals of a portal network} Federal republic, each federal state and each commune must provide an administration portal. Portals of communes must be linked to the portal of the corresponding federal state. Portals of federal republic and federal states must be connected through a portal network. \cite{ozg:onlinegateway_general} Each portal must provide a "seek and find" feature, which enables users to find all administrative services provided by any administration portal \cite{ozg:onlinegateway_architecture}.

\paragraph{Interoperable user profiles for accessing administrative services} Federal republic and federal states must provide user profiles that can be used to identify the corresponding person while requesting access to administrative services, to save personal information according to the once-only principle, to receive and send messages via a digital mailbox and to pay for services. The user profiles must be interoperable for every administration portal of the portal network \cite{ozg:user_profiles_general}.

@@

The implementation of OZG regulations can be separated into two projects: The digitalization and networking project.

The goal of digitalization in this case is to make administrative services available towards a user in a digitized way: A digitized administrative service can be accessed by a user through a website. The website is hosted either by the federal republic or a federal state and is called "Administration Portal". Access is usually provided through an application form. The administrative service can be managed through a user profile which is provided by the federal republic or the federal state. Management of an administrative service usually includes starting the service by sending in a form, communicating with the responsible institutions through the inbox of the profile, and receiving a result. The user profile can also enable users to upload documents to a datawallet and to save personal information for automatically filling in forms.

Digitalization of an administrative service is the modification of the underlying processes to incorporate the usage of the described features of the user profile and administration portal. In total, the BMI lists 575 relevant services, some of them provided by the federal republic, some by the federal states, and some by the communes \cite{BMI:Onlinezugangsgesetz}.

@@

To save investment, a method called "one for all" is used when hosting administrative services. One federal state or the federal republic provides access to a service on their administration portal and distributes the requests to the responsible institutions "under the hood". As administration portals are connected through an "online gateway", each portal contains a search feature, which enables users to find all administrative services through any portal. Interoperable user profiles enable the usage of each profile for the management of administrative services on all portals.

@@

A user wants to apply for two administrative services: funding under the Federal Training Assistance Act (BAföG) and a drivers license. Both administrative services are provided through the "one-for-all" method. The federal state Sachsen-Anhalt is responsible for processing BAföG applications on its administration portal and the federal state Hessen is responsible for processing drivers license applications on its administration portal. The user has never accessed OZG services before.

\paragraph{Create User Profile} The user wants to apply for BAföG first and therefore visits an administration portal which informs that to apply for administrative services, a user profile is required. As user profiles are designed to be interoperable, it does not matter through which administration portal the user profile is created.

The user is requested to enter personal information and to specify an authentication method. For authentication, for example, a username and password combination can be used.

@@

\paragraph{Login to User Profile} Before being able to apply for administrative services, the user has to login into the user profile on the administration portal. Depending on the authentication method used for logging in, the trust level of the current session is determined. Only with a high trust level, users can apply for a broad selection of administrative services. The user has selected an authentication method which results in a high trust level.

\paragraph{Application} On the web page of the administration portal, the user clicks on a "bring me to the application" button and is forwarded to a web page with an interactive form. Before the user is forwarded, he has the option to transfer personal information for automatic filling of the form. The user decides to accept and is presented with a filled-in form. After pressing the "submit" button, the application is transferred to the responsible institution.

\paragraph{Management of Applications} The user is forwarded back to the administration portal and selects to view all active applications created through the portal. Here the user can see the application and is able to manage it by, for example, requesting cancellation.

\paragraph{Communication} Through the inbox of the user profile, the user is notified about a message regarding the application. The institution responsible for processing the application informs the user about receiving the application. Eventually, the institution finishes processing the application and sends a message to the inbox of the user to inform them about the result. This marks the end of the basic OZG use case, however, the user wants to apply for a second administrative service.

\paragraph{Selection of Administrative Service} Using the search feature of the administration portal, the user now enters the search term "drivers license", selects the correct search result, and is forwarded to the corresponding web page on the administration portal of Hessen.

@@

\paragraph{Application} The application process is designed to be similar for each administration portal, however, as Sachsen-Anhalt and Hessen operate their application processes independently, minor differences exist in, for example, the design and text of the web pages.

\paragraph{Management of Applications} The user is again forwarded back to the administration portal and selects to view all active applications created through the portal. The user is only able to see and manage the application for a driving license. As the application for BAföG was created through a the administration portal of Sachsen-Anhalt, it is not accessible through the administration portal of Hessen.

@@

\subsection{System Architecture}
This section describes the OZG system architecture of a federal state capable of executing the basic OZG use case. References about OZG system architectures of Baden-Württemberg \cite{ozg:bw} and Nordrhein-Westfalen \cite{ozg:nrw} are used. The presentation document regarding the administration portal of Baden-Württemberg \cite{ozg:bw_administration_portal} is used as the basis for describing the administration portal.

As shown in figure \ref{ozg:system_overview}, the OZG system architecture is separated into three components and two domains. Domains in this case separate systems regarding the physical location where they operate. Components in this case describe functionally correlated systems of a system architecture.

The "User Component" directly interacts with users to enable access to administration portals. This can be desktop or mobile applications or as in this specific case, a web browser. The "Administration Portal Component" hosts the administration portal. The "User Profile Component" provides interoperable user profiles. The "Integration Component" transfers filled in applications to responsible institutions. And finally, the "Institution Component" receives and processes applications.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{Diagrams/OZG System Overview.pdf}
    \caption{OZG System Architecture of Federal State}
    \label{ozg:system_overview}
\end{figure}

The administration portal directly interacts with the "User Component" through a web page and provides access to the basic OZG use case by interacting with other components. It consists of a web server that is responsible for hosting web pages and a service facade connected to multiple service components. Through the service facade, the web server accesses each service component and makes its service available to the "User Component". 
The "User Profile Management" component is capable of interacting with the "User Profile Component" to create new user profiles, retrieve personal information from existing user profiles, and process login requests. 
The "User Profile Component" provides conventional user profile-based identity management. 
Through the Federated Identity Management of Interoperable User Accounts in Germany (Föderiertes Identitätsmanagement interoperabler Nutzerkonten in Deutschland - FINK) the "User Profile" component of each federal state is made interoperable. The "Search" component can be used to retrieve a list of URLs of administrative services corresponding to a search term. It can retrieve URLs of administrative services from all administration portals through the "Online-Gateway".

The "Online-Gateway" is a system connecting all administration portals and enabling them to exchange information. The "Formula Engine" component is accessed by the web server to display interactive forms to the user when applying for selected administrative services. The interactive form can be either integrated into the web page of the administrative service or hosted on a separate web page. The user can opt in to automatically fill in forms with personal information. In that case, after authentication of the user, the "Administration Portal" component retrieves personal information from the "User Profile" component and sends it to the "Formula Engine" before displaying a filled-in form to the user.

The "Application Management" component receives submitted applications by the "Formula Engine" component for further processing. Depending on the administration portal and the type of application, additional processing steps that are not part of the use case can occur. This includes, for example, payment processes. The component tracks the current status of the application, which the user can view through the web page of the administration portal. It is also capable of terminating the application process if the user requests it.

### 2021-03-26 chapters/sections/7-messaging.tex
finished

This chapter gives a brief overview of what messaging is and how it can be used for system integration purposes. It introduces various messaging patterns which will be used by the technological integration in chapter 4 and 5.
The content of this chapter is based on the book \textit{Enterprise Integration Patterns} by \textcite{EIP}.

Messaging is an asynchronous system-to-system communication, delivering packages of data called messages through channels. Messages can contain any structured data like strings, arrays or objects. It is up to the sender and receiver to determine and understand its content.

Channels can be seen as a queue of messages, accessible to interested systems. Channels have a direction in a sense so that no system simultaneously writes to and reads from a channel. Therefore, by one system writing messages to a channel and another system reading them, a message and data flow can be established.

@@

During the routing of messages between systems, the content of messages might have to be modified. Therefore various message transformation components exist. Their purpose is to consume messages from an inbound channel, modify their content based on predefined or dynamic rules and publish the modified message to an outbound channel.

"\textit{A messaging system manages messaging the way a database system manages data persistence}" \cite[p. 31]{EIP}. Configuring a messaging system involves definition of available channels, the type of messages they exchange, their connection between systems, the message flow and when necessary the insertion of routing and transformation components.

Depending on the messaging system solution, additional features can be configured to increase performance, stability, failure safety and more. If properly configured, the messaging system operates reliably, handles unreliable network connections and persistently stores messages in case of a system crash.

@@

As channels are able to function as a message buffer, sender and receiver of messages do not even have to be running at the same time.

@@

In the book "\textit{Enterprise Integration Patterns}" by \cite{EIP} multiple patterns for designing messaging systems are presented. This section gives a brief overview of patterns relevant for the thesis and their visual notation. Lucidchart \cite{lucid} was used for drawing messaging figures.

@@

There are different types of channels. In this thesis, named channels are publish-subscribe datatype channels. Unnamed channels are point-to-point channels. Publish-subscribe channels can be accessed by every system. In contrast to that, point-to-point channels can only be accessed by two systems - one writing and one reading system.

@@

Systems that do not natively support message based communication require a messaging adapter which enables the programs of the system to access the messaging system through an API. The messaging adapter usually has to be developed individually for each system, and manually integrated into the code of the system through asynchronous function calls of the adapter API.

@@

Content-Based Routers contain a preconfigured set of routing rules, that instruct the router to place inbound messages on different outbound channels depending on their content. These rules do not change during operation of the system.

@@

Message Translators have the purpose of translating message content on data representation and data type layer. 
On the data representation layer, message translators translate message content for example from an XML format to a JSON format. On the data type layer, message translators can for example concatenate first name and last name fields to a single name field.

@@

As shown in figure \ref{messaging:filter2}, Message Filters can, in combination with publish-subscribe channels, be used to separate messages onto individual datatype channels.

@@

Message Stores have the purpose of storing consumed messages in a database. They are also capable of communicating with other messaging patterns for providing access to stored messages. Communication between message patterns and Message Stores is indicated by a dotted arrow.

Figure \ref{messaging:store} shows the visual notation of a Message Store. It also shows the example of a Content Enricher accessing the message store.

>>>>

This chapter gives a brief overview of what messaging is and how it can be used for system integration purposes. It introduces various messaging patterns which will be used by the technological integration in chapters 4 and 5.
The content of this chapter is based on the book \textit{Enterprise Integration Patterns} by \textcite{EIP}.

Messaging is an asynchronous system-to-system communication, delivering packages of data called messages through channels. Messages can contain any structured data like strings, arrays, or objects. It is up to the sender and receiver to determine and understand its content.

Channels can be seen as a queue of messages accessible to interested systems. Channels have a direction in a sense so that no system simultaneously writes to and reads from a channel. Therefore, by one system writing messages on a channel and another system reading them, a message and data flow can be established.

@@

During the routing of messages between systems, the content of messages might have to be modified. Therefore, various message transformation components exist. Their purpose is to consume messages from an inbound channel, modify their content based on predefined or dynamic rules, and publish the modified message to an outbound channel.

A messaging system manages messaging the way a database system manages data persistence \cite[p. 31]{EIP}. Configuring a messaging system involves the definition of available channels, the type of messages they exchange, their connection between systems, the message flow, and, when necessary, the insertion of routing and transformation components.

Depending on the messaging system solution, additional features can be configured to increase performance, stability, failure safety and more. If properly configured, the messaging system operates reliably, handles unreliable network connections, and persistently stores messages in case of a system crash.

@@

As channels are able to function as a message buffer, the sender and receiver of messages do not even have to be running at the same time.

@@

In the book "\textit{Enterprise Integration Patterns}" by \textcite{EIP} multiple patterns for designing messaging systems are presented. This section gives a brief overview of patterns relevant for the thesis and their visual notation. Lucidchart \cite{lucid} was used for drawing messaging figures.

@@

There are different types of channels. In this thesis, the named channels are publish-subscribe datatype channels. Unnamed channels are point-to-point channels. Publish-subscribe channels can be accessed by every system. In contrast to that, point-to-point channels can only be accessed by two systems - one writing and one reading system.

@@

Systems that do not natively support message-based communication require a messaging adapter which enables the programs of the system to access the messaging system through an API. The messaging adapter usually has to be developed individually for each system, and manually integrated into the code of the system through asynchronous function calls of the adapter API.

@@

Content-Based Routers contain a preconfigured set of routing rules that instruct the router to place inbound messages on different outbound channels depending on their content. These rules do not change during operation of the system.

@@

Message Translators have the purpose of translating message content on data representation and data type layer. 
On the data representation layer, message translators translate message content, for example, from an XML format to a JSON format. On the data type layer, message translators can, for example, concatenate first name and last name fields to a single name field.

@@

As shown in figure \ref{messaging:filter2}, Message Filters can, in combination with publish-subscribe channels, be used to separate messages to individual datatype channels.

@@

Message Stores have the purpose of storing consumed messages in a database. They are also capable of communicating with other messaging patterns to provide access to stored messages. Communication between message patterns and Message Stores is indicated by a dotted arrow.

Figure \ref{messaging:store} shows the visual notation of a Message Store. It also shows an example of a Content Enricher accessing the message store.

### 2021-03-26 chapters/sections/8-identity_management.tex
Einleitung überarbeitet
finished

This chapter describes issues of user based identity management in general and for the specific use case of the OZG. It also presents possibilities of how IMP can be utilized to solve mentioned issues.

@@

The OZG system architecture utilizes user profiles for identity management. This leads to several issues regarding usability, data protection, security and maintenance costs. The following sections describe problems of user profiles in general as well as problems arising specifically in context of the OZG system architecture.

\paragraph{Usability}
Each Service Provider has a different platform for the management of personal information and identity related interactions. This can for example be web pages or applications for desktop and smartphone. Keeping track of all created user profiles is difficult.

In case of web pages, users will also have to remember various URLs, usernames and passwords. The risk of forgetting login credentials is high.

To manage personal information stored in each user profile and for example change an e-mail address, users have to access the individual platforms of each Service Provider. Depending on the amount of created profiles, this can be a time consuming task. Especially if the user has to login to web pages using various authentication methods like passwords, single-sign-on, 2FA or Magic-Link. Users receiving notifications, messages and mails on multiple platforms can quickly loose track and oversee important notifications.

Each platform is often presented with a different design although the purpose remains similar. This makes it difficult for users to orientate themselves.

Bidirectional identity related interactions like subscribing to a service, canceling a subscription, sending or receiving a security relevant mail often only takes place through a web page where e-mails are only used to notify the user to check his user profile ("You have unread messages!"). This is because e-mails do not provide a secure way of exchanging information and do not enable advanced interaction capabilities or exchange of structured content. Although Service Providers could develop mobile applications as an alternative for the web pages and benefit from secure communication and advanced interaction capabilities, due to the required investments many smaller SPs decide not to.

Service Providers rely on the validity of information stored in the user profile. To guarantee that the information is correct, some SPs require users to take extra identification steps by for example performing a video call or requesting the user to show their ID at a local shop. As each Service Provider manages separate user profiles, the user might have to go through this time consuming process repeatedly.

@@

In case of the OZG, usability of user profiles across multiple administration portals is limited: Although user profiles are interoperable, administration portals are individually responsible for the application process. Applying for different administration services, users might have to visit separate administration portals. Each of the portals manage the application process differently. Active applications are not stored as part of the user profile but remain accessible only through the respective portals they were created through.\\ An overview of all active applications amongst all administration portals does not exist. The user has to go through each portal to find the application they for example want to cancel. Especially due to the „one for all“ approach, users will have to access dedicated administration portals for certain types of administrative services. To mitigate this issue, the inbox of user profiles is used to notify the user of the status of all applications.

\paragraph{Data Protection}
Before, Service Providers established a habit of collecting more personal information than necessary to provide their services. As a response, the "Datenschutz Grundverordnung" - DSGVO (General Data Protection Regulation) was passed. Amongst other things, it requires Service Providers to inform the user about which personal information is processed, the reason for processing the information and how it is being processed. Conventional user profile systems which store all personal information but do not differentiate between possibly different processing legitimisation of attributes increase the risk of illegitimate data usage.

@@

In case of the OZG, the user shares personal information not only with institutions responsible for processing an application but also with multiple systems operated by the member state. Especially the administration portal used to initiate the application is problematic as it receives personal information from the interoperable user profile, sends it to a form server, eventually receives the filled in form and stores it. As a result of that, the administration portal has access to all personal information that in reality only the institution will eventually need to process. In addition, institutions actually do not require user profiles for processing applications. After an application is finished, the institution can delete all data stored as part of this process. The only reason user profiles are required is to verify the validity and authorization of incoming applications and to enable the user to comfortably fill in application forms and communicate through an inbox.

@@

User profiles are most commonly secured through passwords. Especially in case of a large amount of profiles, users are tempted to use identical passwords, increasing the risk of them being stolen and all user profiles being accessible. With the amount of different Service Providers storing personal information and passwords, the risk of it being stolen increases as security of system architectures differs between Service Providers.

Users also have to rely on the correct operation of the system architecture of the Service Provider and might not be able to prove failures in the operation of the system. If for example the user cancels a subscription but the system architecture quietly fails to process it, the user might have issues proving to the Service Provider, that they did cancel the subscription.

@@

The goal of IMP solutions is to replace user profiles with a single IMP identity. Through the IMP client a user can manage their IMP identity and all connected integrated Service Providers. Instead of creating user profiles, the user can establish a relationship between their IMP identity and the Service Provider. Depending on the use case, the relationship can be utilized differently. Using the IMP client, the user can keep track of all relationships without the need of remembering URLs, usernames or passwords. It also enables the user to update personal information shared with Service Providers through one application. As the IMP client also contains an inbox, all notifications sent by each Service Provider can be accessed in the same place.

As part of established relationships, IMP client and Service Provider can securely exchange bidirectional structured data which along with advanced interaction capabilities of the IMP client can be used to create custom interaction scenarios: Service Providers can send structured data to the IMP client along with the information that it should display as a form. Depending on the content of the form and the use case of the Service Provider, the form can for example be a purchase agreement or a survey.

Assuming that, through a secure identification and validation process, the identity management provider has made sure that attributes users store as part of their IMP identity are correct, Service Providers wont require any additional identification steps through for example video calls.

Due to the IMP client being able to automatically fill in attributes requested as part of a relationship, the user usually wont need to repeatedly manually enter personal information.

In case of the OZG system architecture, IMP can be used by each member state to individually establish relationships with IMP identities of users to access personal information. This could replace the requirement for user profiles to be interoperable. For the most part, member states would then be able to operate independently from each other. This solution is described in detail in chapter 4.

@@

IMP puts the users back in control of their personal information. \\
As part of the process of establishing a relationship, the Service Provider has to specify exactly which attributes have to be shared, give a detailed explanation of why the attributes are required to be shared, what the purpose of the relationship is and how the attributes are processed. With this information, the user can make an informed decision whether to establish the relationship and the Service Provider performed their duty to inform the user as required by the DSGVO.

@@

As every relationship along with shared attributes is manageable through the IMP client, the user wont loose track of which Service Provider has access to what personal information.

The risk of data being stolen decreases as most of the identity management and data storage is performed by the IMP system and the Service Provider processes only a small portion of attributes of IMP identities.

In case of the OZG system architecture, IMP can be used by the individual institutions to establish relationships with IMP identities of users to retrieve applications for administrative services. This could eventually eliminate the need for user profiles. Personal information could be shared with only the responsible institution. This solution is described in detail in chapter 5.

@@

The IMP client enables users to access their IMP identity and all relationships without the need of remembering a variety of credentials. This removes the risk of credentials being stolen. The IMP client itself can be secured without the need for a password. As the IMP system is capable of signing relationships and exchanged data, interactions between IMP identity and Service Provider can be validated at any time. This enables both user and Service Provider to provide proof for every interactions.

\section{Outcome}

>>>>

This chapter describes the issues of user-based identity management in general and for the specific use case of the OZG. It also presents possibilities of how IMP can be utilized to solve the mentioned issues. At the end, the conclusion briefly states whether IMP can be considered for solving the problems of user profile-based identity management.

@@

The OZG system architecture utilizes user profiles for identity management. This leads to several issues regarding usability, data protection, security, and maintenance costs. The following sections describe problems of user profiles in general as well as problems arising specifically in the context of the OZG.

\paragraph{Usability}
Each Service Provider has a different platform for the management of personal information and identity-related interactions. This can, for example, be web pages or applications for personal computers and smartphones. Keeping track of all created user profiles is difficult.

In the case of web pages, users will also have to remember various URLs, usernames, and passwords. The risk of forgetting login credentials is high.

To manage the personal information stored in each user profile and, for example, change an e-mail address, users have to access the individual platforms of each Service Provider. Depending on the amount of created profiles, this can be a time-consuming task. Especially if the user has to login to web pages using various authentication methods like passwords, single-sign-on, 2FA, or Magic-Link. Users receiving notifications on multiple platforms can quickly lose track and oversee important notifications.

Each platform is often presented with a different design, although the purpose remains similar. This makes it difficult for users to orientate themselves.

Bidirectional identity-related interactions like subscribing to a service, canceling a subscription, sending or receiving a security relevant mail often only takes place through a web page where e-mails are only used to notify the user to check his user profile ("You have unread messages!"). This is because e-mails do not provide a secure way of exchanging information and do not enable advanced interaction capabilities or exchange of structured content. Although Service Providers could develop mobile applications as an alternative for web pages and benefit from secure communication and advanced interaction capabilities, due to the required investments, smaller SPs might decide not to.

Service Providers rely on the validity of the information stored in the user profile. To guarantee that the information is correct, some SPs require users to take extra identification steps by, for example, performing a video call or requesting the user to show their ID at a local shop. As each Service Provider manages separate user profiles, the user might have to go through this time-consuming process repeatedly.

@@

In the case of the OZG, the usability of user profiles across multiple administration portals is limited: Although user profiles are interoperable, the administration portals are individually responsible for the application process. Applying for different administration services, users might have to visit separate administration portals. Each of the portals manages the application process differently. Active applications are not stored as part of the user profile but remain accessible only through the respective portals they were created through.\\ An overview of all active applications amongst all administration portals does not exist. The user has to go through each portal to find the application they, for example, want to cancel. Especially due to the „one-for-all“ approach users will have to access dedicated administration portals for certain types of administrative services. To mitigate this issue, the inbox of user profiles is used to notify the user of the status of all applications.

\paragraph{Data Protection}
Service Providers established a habit of collecting more personal information than necessary to provide their services. As a response, the "Datenschutz Grundverordnung" - DSGVO (General Data Protection Regulation) was passed. Amongst other things, it requires Service Providers to inform the user about which personal information is processed, the reason for processing the information, and how it is being processed. Conventional user profile systems which store all personal information but do not differentiate between possibly different processing legitimisation of attributes increase the risk of illegitimate data usage.

@@

In the case of the OZG, the user shares personal information not only with the institutions responsible for processing an application but also with multiple systems operated by the member state. Especially the administration portal used to initiate the application is problematic as it receives personal information from the interoperable user profile, sends it to a form server, eventually receives the filled-in form and stores it. As a result of that, the administration portal has access to all personal information that in reality only the institution will eventually need to process. In addition, institutions actually do not require user profiles for processing applications. After an application is finished, the institution can delete all data stored as part of this process. The only reason user profiles are required is to verify the validity and authorization of incoming applications and to enable the user to comfortably fill in application forms and communicate through an inbox.

@@

User profiles are often secured through passwords. Especially for a large number of profiles, users are tempted to use identical passwords, increasing the risk of them being stolen and all user profiles being accessible. With the number of different Service Providers storing personal information and passwords, the risk of it being stolen increases as the security of system architectures differs between Service Providers.

Users also have to rely on the correct operation of the system architecture of the Service Provider and might not be able to prove failures in the operation of the system. If, for example, the user cancels a subscription but the system architecture quietly fails to process it, the user might have issues to provide proof.

@@

The goal of IMP solutions is to replace user profiles with a single IMP identity. Through the IMP client, a user can manage their IMP identity and all connected integrated Service Providers. Instead of creating user profiles, the user can establish a relationship between their IMP identity and the Service Provider. Depending on the use case, the relationship can be utilized differently. Using the IMP client, the user can keep track of all relationships without the need of remembering URLs, usernames, or passwords. It also enables the user to update personal information shared with Service Providers through one application. As the IMP client also contains an inbox, all notifications sent by each Service Provider can be accessed in the same place.

As part of established relationships, IMP clients and Service Providers can securely exchange bidirectional structured data, which along with the advanced interaction capabilities of the IMP client, can be used to create custom interaction scenarios: Service Providers can send structured data to the IMP client along with the information that it should display as a form. Depending on the content of the form and the use case of the Service Provider, the form can, for example, be a purchase agreement or a survey.

Assuming that, through a secure identification and validation process, the identity management provider has made sure that the attributes users store as part of their IMP identity are correct, Service Providers wont require any additional identification steps through for example video calls.

Due to the IMP client being able to automatically fill in the attributes requested as part of a relationship, the user usually wont need to repeatedly manually enter personal information.

In the case of the OZG system architecture, IMP can be used by each member state to individually establish relationships with IMP identities of users to access personal information. This could replace the requirement for user profiles to be interoperable. For the most part, the member states would then be able to operate independently from each other. This solution is described in detail in chapter 4.

@@

IMP puts the users back in control of their personal information. \\
As part of the process of establishing a relationship, the Service Provider has to specify exactly which attributes have to be shared, give a detailed explanation of why the attributes are required to be shared, what the purpose of the relationship is and how the attributes are processed. The user can make an informed decision whether to establish the relationship and the Service Provider performs its duty to inform the user as required by the DSGVO.

@@

As every relationship is manageable through the IMP client, the user wont loose track of which Service Provider has access to which personal information.

The risk of data being stolen decreases as most of the identity management and data storage is performed by the IMP system and the Service Provider processes only a small portion of the attributes of IMP identities.

In the case of the OZG system architecture, IMP can be used by individual institutions to establish relationships with IMP identities of users to retrieve applications for administrative services. This could eventually eliminate the need for user profiles. Personal information could be shared with only the responsible institution. This solution is described in detail in chapter 5.

@@

The IMP client enables users to access their IMP identity and all relationships without the need of remembering a variety of credentials. This removes the risk of credentials being stolen. The IMP client itself can be secured without the need for a password. As the IMP system is capable of signing relationships and exchanged data, the interactions between IMP identity and Service Provider can be validated at any time. This enables both the user and Service Provider to provide proof for every interaction.

\section{Conclusion}

User profile based identity management especially in case of the OZG has several problems in regard to usability, data protection and security. As described, in many cases, IMP has the possibility of solving these issues and eliminating or at least enabling better control over fragmented partial identities. This is the case for interoperable OZG user profiles as well as user profiles in general. Therefore it makes sense to further investigate possibilities of utilizing IMP as a solution to partial identities. Using the small scale model of fragmented partial identities in context of the OZG, the following chapters 4 and 5 present different possibilities of utilizing IMP.
